{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory

### 2023-10-29 paper/main.tex
minor changes after merge
fix reward notation for consistency with si

In fact, even beyond specific psychological considerations, a curious interpretation of payoffs arises from the formula commonly used for expected payoffs in repeated games. If $\delta\in\left[0,1\right)$ is the probability of continuing to another round in the game, then the expected payoff to an agent is $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$, where $\pi_{t}$ is the payoff the agent receives in the stage game at time $t$. Here, additional stochasticity arises due to uncertainty in the game length, and an agent might not be able to compute his or her expected payoff for use in decision-making. The probability that the game terminates after the interaction at time $T$ is $\left(1-\delta\right)\delta^{T}$, in which case $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$ is exactly the expected payoff the agent receives at time $T$, i.e. \emph{in the last round}. As an unbiased estimator of this expectation, the agent might thus use $\pi_{T}$ as a proxy for ``success'' when evaluating his or her behavior. This gives a purely model-driven justification for why considering payoff in the last round of the game can result in more realistic extensions of traditional models.

We note that the expected \emph{total} payoff in the game, $\pi_{0}+\pi_{1}+\cdots +\pi_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$. This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}\pi_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$). Non-normalized payoffs interfere with selection intensity ($\beta$) in models of social imitation, which is (presumably) why they appear less frequently in the literature. On that point, we note that many of the differences between realized and expected payoffs disappear in the limit of weak selection \citep{mcavoy:NHB:2020}. Non-weak selection can introduce substantial differences between models with realized and expected payoffs \citep{mcavoy:PLOSCB:2021}, which is especially important to understand in models of social systems with cultural transmission \citep{cavalli1981cultural}.

@@

{
{\setlength{\bibsep}{0\baselineskip}
\bibliographystyle{naturemag}
\bibliography{bibliography.bib}
}
}

>>>>

In fact, even beyond specific psychological considerations, a curious interpretation of payoffs arises from the formula commonly used for expected payoffs in repeated games. If $\delta\in\left[0,1\right)$ is the probability of continuing to another round in the game, then the expected payoff to an agent is $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$, where $r_{t}$ is the reward the agent receives in the stage game at time $t$. Here, additional stochasticity arises due to uncertainty in the game length, and an agent might not be able to compute his or her expected payoff for use in decision-making. The probability that the game terminates after the interaction at time $T$ is $\left(1-\delta\right)\delta^{T}$, in which case $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$ is exactly the expected payoff the agent receives at time $T$, i.e. \emph{in the last round}. As an unbiased estimator of this expectation, the agent might thus use $r_{T}$ as a proxy for ``success'' when evaluating his or her behavior. This gives a purely model-driven justification for why considering payoff in the last round of the game can result in more realistic extensions of traditional models.

We note that the expected \emph{total} payoff in the game, $r_{0}+r_{1}+\cdots +r_{T}$, is given by $\sum_{t=0}^{\infty}\delta^{t}r_{t}$. This version of ``expected payoff'' appears less common in the literature on direct reciprocity than its normalization, $\left(1-\delta\right)\sum_{t=0}^{\infty}\delta^{t}r_{t}$, likely owing to the fact that payoffs can grow arbitrarily large with sufficiently long time horizons ($\delta\rightarrow 1^{-}$). Non-normalized payoffs interfere with selection intensity ($\beta$) in models of social imitation, which is (presumably) why they appear less frequently in the literature. On that point, we note that many of the differences between realized and expected payoffs disappear in the limit of weak selection \citep{mcavoy:NHB:2020}. Non-weak selection can introduce substantial differences between models with realized and expected payoffs \citep{mcavoy:PLOSCB:2021}, which is especially important to understand in models of social systems with cultural transmission \citep{cavalli1981cultural}.

@@

{
{\setlength{\bibsep}{0\baselineskip}
\bibliographystyle{naturemag}
\bibliography{bibliography}
}
